Paper Report

### 2017-10-18 Paper_report.tex
Unupervised Neurall HMM
Unsupervised Neural HMM

\section{Unsupervised Neural HMMs \citep{Tran16unsupervised}}
\subsection{Main idea}
This research show how to apply unsupervised hidden Markov model in neural network approach.

>>>>

\section{Unsupervised Neural HMMs \citep{Tran16unsupervised}}
\subsection{Main idea}
This research show how to apply unsupervised hidden Markov model in neural network approach. In fact, they would like to prove that a simple nn models trained to maximize the marginal likelihood could outperform more complicated models in unsupervised learning.

\subsection{What is their concentration ?}
\begin{itemize}
\item There are three components:
	\begin{itemize}
	\item Set of latent variables Z (Tags)
	\item Set of observed variables X (Words)
	\item Model parameters $\theta$ (Emission and transitions probability)
	\end{itemize}

\item Purpose: Find $\theta$ which maximize p(X $|\theta$)
\item How: Use Generalized EM to estimate $\theta$
	\begin{enumerate}
	\item : Maximizing $p(X)$ means
	\begin{equation}
	p(x) = \sum_{z} p(X,Z) = E_{q(Z)}[\ln p(X,Z|\theta)] + H[q(Z)] + KL(q(Z) || p(Z|X,\theta))
	\end{equation}
	\item E-step: Estimate $p(Z|X)$ based on current $\theta$
	\item M-step: Consider $q(Z) = p(Z|X)$ \\
	$KL(q(Z) || p(Z|X,\theta)) = 0$ \\
	$H[q(Z)]$ constant
	\item Maximizing $p(X)$ becomes maximizing $E_{q(Z)}[\ln p(X,Z|\theta)]$
	\end{enumerate}
\item Result: Gradient of the joint probability scaled by the posteriors
	\begin{equation}
	J(\theta) = \sum_{Z} p(Z|X) \dfrac{\partial}{\partial\theta} \ln p(X,Z| \theta)
	\label{GradientJointProbability} 
	\end{equation}
\item Problem: How to calculate $p(X,Z)$ ?
\end{itemize}

\subsection{What is the role of HMM ?}
\begin{itemize}
\item Assumption:
	\begin{itemize}
	\item Every word token is generated by a latent class (Tag)
	\item The current class at time t is conditioned on the previous class at time (t - 1)
	\end{itemize}

\item Therefore, the probability of a given sequence of observation X and latent variables Z (Factorization of the joint probability):
	\begin{equation}
	p(X,Z) = \prod_{t=1}^{n + 1} p(z_{t}|z_{t-1})\prod_{t=1}^{n} p(x_{t}|z_{t})
	\label{ProbabilityHMM}
	\end{equation}

\item Result: Combine $p(X,Z)$ \eqref{ProbabilityHMM} and gradient $J(\theta)$ \eqref{GradientJointProbability}

$J(\theta) = \sum_{Z} p(Z|X) \dfrac{\partial}{\partial\theta} \ln p(X,Z| \theta)$ \\

$ = \sum_{Z} p(Z|X) \dfrac{\partial}{\partial\theta} \ln [\prod_{t=1}^{n + 1} p(z_{t}|z_{t-1},\theta)\prod_{t=1}^{n} p(x_{t}|z_{t},\theta) ]$ \\

$ = \sum_{Z} p(Z|X) \dfrac{\partial}{\partial\theta}  [\sum_{t=1}^{n + 1} \ln p(z_{t}|z_{t-1},\theta) + \sum_{t=1}^{n} \ln p(x_{t}|z_{t},\theta) ]$

$ = \sum_{t}\sum_{z_{t}} p(z_{t}, z_{t-1}|X) \dfrac{\partial}{\partial\theta} \ln p(z_{t}|z_{t-1},\theta) + p(z_{t}|X) \dfrac{\partial}{\partial\theta} \ln p(x_{t}|z_{t},\theta) $

\begin{equation}
	J(\theta) = \sum_{t}\sum_{z_{t}} p(z_{t}, z_{t-1}|X) \dfrac{\partial}{\partial\theta} \ln p(z_{t}|z_{t-1},\theta) + p(z_{t}|X) \dfrac{\partial}{\partial\theta} \ln p(x_{t}|z_{t},\theta)
\end{equation}

\item Problem:
	\begin{itemize}
	\item How to calculate $p(z_{t}, z_{t-1}|X)$ and $p(z_{t}|X)$ ? They propose Baum-Welch
	\item How to calculate $p(z_{t}|z_{t-1},\theta)$ and $p(x_{t}|z_{t},\theta)$ ? They propose Neural Networks
	\end{itemize}
\end{itemize}

\subsection{Where is neural network ?}
\begin{itemize}
\item Input: A sentence $X = x_{1},..., x_{t},... x_{L_{x}}$, a set of vocabulary $W = w_{1},..., w_{i},... w_{L_{W}}$ a set of tags $Z = z_{1},..., z_{j},... z_{L_{z}}$
\item Output: $p(z_{t}|z_{t-1},\theta)$ and $p(x_{t}|z_{t},\theta)$ at each time t.
\item How:
	\begin{enumerate}
	\item Embedding X and Z by $\theta$: Vector embedding of W $v_{W}$ (Using CNN - Convolution for Morphology) and vector embedding of Z $v_{Z}$ (Simple feed-forward nn having a lookup table following by a non-linear activation ReLU). $v_{W}$ and $v_{Z}$ have the same dimension.
	\item Calculate $p(x_{t}|z_{t},\theta)$ (Emission matrix): Probability of a word $w_{i}$ in Vocabulary is generated by a tag $z_{j}$ (Do not care about time t).
	\begin{equation}
	p(w_{i}|z_{j}) = \frac{exp(v_{z_{j}}^{T}*v_{w_{i}} + b_{i})}{\sum_{w \in W} exp(v_{z_{j}}^{T}*v_{w} + b)}
	\end{equation}

	\item Calculate $p(z_{t}|z_{t-1},\theta)$ (Transition matrix): Probability of a tag $z_{j}$ at time t is generated by a tag $z_{j'}$ at time (t - 1).\\
	Input: Vector of word w at $x_{t}$ noted $v_{x_{t}}$. It is query embedding (Using LSTMs) \\
	Result: Matrix of $L_{z}*L_{z}$ noted T. It means all transition probabilities of each tag at (t - 1) to all tags at time t.\\
	\begin{equation}
	T = U^{T}*v_{x_{t}} + b
	\end{equation}
	\end{enumerate}
\end{itemize}
